Designing Turk Experiments for Visual Statistical Inference

### 2013-07-06 paper_turk.tex
reviwerd upto the end of section 2.3 (sample size estimatin)

Human observers are needed to evaluate lineups that are used to test the significance of findings using statistical graphics. One good option is to recruit people from online workplace like Amazon Mechanical Turk (MTurk). It provides features to create online task that allow people to evaluate lineups and get paid. MTurk is designed for simple and easy tasks. The technical design of the underlying experiment for lineup evaluation may be complex and the tools available to design this from MTurk is just too simple. In this paper we present the design of MTurk experiments for lineup evaluation, provide an alternative way to conduct the survey on lineups by developing a separate web application and getting turk worker do their job from that web site. The web site is now hosted on the Iowa State University public domain and has been in use for multiple experiments \citep{majumder:turk}. It provides multiple features that makes it convenient to get lineups evaluated by online observers and obtain data in a secure way.

@@

There have been some advancements in visual statistical inference since it was first introduced by \cite{buja:2009}. They  proposed lineup protocol that allows testing the significance of findings using statistical graphics. In visual statistical inference the test statistics is a plot of the observed data. This plot is placed randomly in a layout of plots called lineup. The rest of the plots, called null plots, in the lineups are generated from the model specified by the null hypothesis. A human observer is asked to evaluate the lineup. If the observer can detect the actual plot in the lineup, the null hypothesis is rejected.

A lineup is shown in Figure~\ref{fig:lineup_turk} where the observed scatterplot is displayed with a least square line overlaid. Can you find which of these plots show the steepest slope? The null plots are generated based on the hypothesis that the slope is zero. If the observed scatterplot is detected that indicates that observed plot is different from the null plots and hence the null hypothesis is rejected.

@@

\cite{majumder:2013}  further developed visual inference by refining the terminologies and providing the methods of obtaining the power of the visual test. it is revealed that visual test power can be as good as the best available conventional test and in some scenarios even better. This work establishes the validity of lineup protocol to be used as a tool for statistical testing. In the situations where no conventional test exists, visual inference can be the only inferential procedure without compromising a lot in power because of its non-parametric nature and few assumptions.

@@

The simplicity of the MTurk task is the main factor for its popularity. Figure \ref{fig:amazon_task} shows how simple an MTurk task could be. It is possible to get a lineup evaluated by creating such a simple task. We just need to replace the picture in Figure \ref{fig:amazon_task} with the lineup in Figure \ref{fig:lineup_turk} and change the answering options. But some times we may need more than one lineups to be evaluated by an observer. We may need to show a random sample of lineups from a pool of many lineups automatically. The questions the observer would answer while examining the lineup can be different based on different lineup. Moreover, It is convenient if the lineup is clickable so that selection of plot can be made by mouse click. To create an MTurk task that has all these flexibilities, a web application needs to be created. The following section discusses how the system would work.

@@

Once a web application is created, there are two options to run it; one is to run it inside MTurk system using their API and the other is to develop a new web site separate from MTurk. For additional control we picked the second option and planned to separate this application from MTurk system and designed our own web site to run the application. It enables to display the lineups to the observers with a lot of flexibilities. As an added benefit the data can be directly saved in a local database server instead of getting it from MTurk.

Figure \ref{fig:turk_work_flow} shows how the plan works. First an MTurk task is created for the workers to see and decide whether they want to do the task based on the parameters like payment amount, estimated time the task may take to finish, how hard the task seems to be etc. Workers are informed that the task has to be done outside the MTurk system from another web site. Once the workers accept the task they are redirected to our web site where multiple lineups are shown for evaluations. After the required number of evaluations has been received, a code is generated which the workers has to submit back in MTurk system to complete the task. The code is matched with the code in our database to process payment through MTurk system.

@@

If a lineup is created from actual observed data, the web application presented in Section \ref{sec:web_application} can be used to evaluate it. In that case we do not need to design a simulation experiment. This section will be useful when a simulation experiment is needed to examine different visual test statistics and compare the power \citep{majumder:2013}. To examine the effectiveness of certain plots in displaying the data one may want to design an experiment with lineup \citep{heike:2012}. In any cases, two major considerations are choice of parameters to simulate lineup and number of people needed to evaluate a lineup.

\subsection{Selecting Parameter to Simulate Lineup} \label{sec:parameter_selection} In any experimental design we need to have some control. For a simulation experiment with lineup, this refers to the parameter related to a lineup. This depends on what the purpose of the experiment is. For example, to compare the power of a parametric test one may fix the parameter specified by alternative hypothesis and generate data from that model. This data can be considered as the observed data. A lineup can be created by placing this observed data plot in a layout where rest of the data plots come from the model specified by the null model \citep{majumder:2013}.

In addition to the hypothesized parameter of interest, other parameters that may need to be fixed include sample sizes, variability in the data, error structure etc. These parameters produces the effect sizes and are responsible for any detectable signal in the lineup. Thus for a simulation study a range of effects sizes need to be considered. It is a little bit tricky to decide what exact effect sizes are appropriate. For example in \cite{majumder:2013} the parameters were chosen so that they produce a smoother power curve of a conventional test. But in general this depends on what the researchers may want to test and control in an experimental study.

@@

\subsection{Procedure to Simulate Data Plot} \label{sec:simulate_plot} In the simulation study the very first step is to generate a random sample of data from model of interest for pre-specified parameters. We call this data set observed data. Every time we do that we get a new set of observed data which may produce a parameter very different than the actual parameter specified to generate the data. But it is important to have a observed data closely representing the parameters chosen since other null plots will be compared to this data plot in a lineup. One solution to this is to take some replications of lineups for the same effect size.

While the effect of the natural variability of the observed data set can be controlled by taking the replication of few samples, it is desirable to study whether we can reduce this variability further to some extent. For this we pick the example of a simple linear regression model and examines how the data plot can be generated that best represent the specified parameter.  Consider a linear regression model 
\begin{equation} \label{eqn:turk_model} 
Y_i = \beta_0 + \beta X  + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariate $X$ can be continuous or discrete. As discussed in Section \ref{sec:parameter_selection} we specify the parameters sample size $n$=300, regression slope $\beta$=3 and error standard deviation $\sigma$=12 to simulate observed data.

@@

{\bf Quantiles of $p$-value approach:} For a simulated observed data set we can obtain the $p$-value associated to testing $H_0: \beta=0$. The distribution of $p$-value is uniform under null hypothesis but under alternative it has a right skewed distribution. Figure~\ref{fig:dist_pvalue} shows the distribution  of $p$-values for sample size $n$=300, regression slope $\beta$=3 and error standard deviation $\sigma$=12. If we generate a data set randomly there is a 21\% chance that the data will show a $p$-value of 0.25 or more even though we generated data set with non-zero $\beta=3$. We need to make sure that the simulated data does not come from this extreme end. Additionally, we want some replications so that this extreme effect can be controlled at some level. For this example we intend to pick three replications.

\begin{figure}[hbtp]
   \centering
       \scalebox{.6}{\includegraphics{dist_pvalue.pdf}}
       \caption{Distribution of $p$-values for $n$=300, $\beta$=3 and $\sigma$=12.}
       \label{fig:dist_pvalue}
\end{figure}

@@

{\bf Closeness of estimated parameters:} When we simulate a data set from the Model \eqref{eqn:turk_model} and fit the model to the data set again, we do not get the parameters estimates same as what we fixed while simulation. For example, we obtained a parameter estimates of 0.5 for true $\beta=3$. We want a data set where the parameter is closed to our fixed slope $\beta=3$.  In this approach we generate 1000 data sets and pick the data set that shows most close estimates of parameters compared to true parameter values. We do this three times to obtain three replications of the data.

@@

\subsection{Sample size estimation} For any experiment it is important to determine the sample size. It is not only related to time and cost of the experiment but also the validity of the results depends on it. In simulation experiment with lineup the sample size means the number of evaluations per lineup. It is different than the number of people who evaluates lineup. If multiple lineups are evaluated by a person, sample size will be larger than number of people recruited from MTurk. For example, if a lineup needs to be evaluated 20 times, we need at least 20 persons since each person does not see the same lineup more than once. This will also enable us to get 10 lineups evaluated if all of these 20 people evaluate 10 lineups each. But all the lineups don't require same number of evaluations. For easy or difficult lineups fewer evaluations are needed.

One approach to assess number of evaluations that are required for a lineup is to have a prior idea of proportion correct for the that lineup. For a given proportion $p$ we want to have margin of error (ME) to be 0.1. Thus we have $$ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$$ which gives us the estimation of minimum sample size $$n \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$

If we do not have any prior idea of proportion of correct or power, we may rely on the power of a similar conventional test if available. This can be assumed to be the power or proportion correct for that lineup. Other ways of having an estimate of sample size can be very specific to the problem of interest and should be dealt with the requirement of the problem.

>>>>

Human observers are needed to evaluate lineups that are used to test the significance of findings using statistical graphics. One good option is to recruit people from online workplace like Amazon Mechanical Turk (MTurk). It provides features to create online task that allow people to evaluate lineups and get paid. MTurk is designed for simple and easy tasks. The technical design of the underlying experiment for lineup evaluation may be complex and the tools available to design this from MTurk is just too simple. In this paper we present the design of MTurk experiments for lineup evaluation, provide an alternative way to conduct the survey on lineups by developing a separate web application and getting turk worker do their job from that web site. The web site is now hosted on the Iowa State University public domain and has been in use for multiple experiments \citep{majumder:turk}. It provides multiple features that make it convenient to get lineups evaluated by online observers and obtain data in a secure way.

@@

There have been some advancements in visual statistical inference since the concept was first introduced by \cite{buja:2009}. They  proposed lineup protocol that allows testing the significance of findings using statistical graphics. In visual statistical inference the test statistics is a plot of the observed data. This plot is placed randomly in a layout of plots called lineup. The rest of the plots, called null plots, in the lineups are generated using the data simulated from the model specified by the null hypothesis. A human observer is asked to evaluate the lineup. If the observer can detect the actual plot in the lineup, the null hypothesis is rejected.

A lineup is shown in Figure~\ref{fig:lineup_turk} where the observed scatterplot is displayed with a least square line overlaid. Can you find which of these plots show the steepest slope? The null plots are generated based on the hypothesis that the slope is zero. If the observed scatterplot is detected, it indicates that observed plot is different from the null plots and hence provides evidence against the null hypothesis.

@@

\cite{majumder:2013}  further developed visual statistical inference by refining the terminologies, presenting the ways to compute $p$-values associated with the visual test and providing the methods of obtaining the power of the test. It is revealed that the power of visual test  can be as good as that of the best available conventional test and in some scenarios even better. This work establishes the validity of lineup protocol to be used as a tool for statistical testing. In the situations where no conventional test exists, visual inference can be the only inferential procedure that does not compromise a lot on power because of its non-parametric nature and few assumptions.

@@

The simplicity of the MTurk task is the main factor for its popularity. Figure \ref{fig:amazon_task} shows how simple an MTurk task could be. It is possible to get a lineup evaluated by creating such a simple task. We just need to replace the picture in Figure \ref{fig:amazon_task} with the lineup in Figure \ref{fig:lineup_turk} and change the answering options. But some times we may need more than one lineups to be evaluated by an observer. We may need to show a random sample of lineups from a pool of many lineups automatically. The questions the observer would answer while examining the lineup can be different based on different lineups. Moreover, It is convenient if the lineup is clickable so that selection of plot can be made by the mouse click. To create an MTurk task that has all these flexibilities, a web application needs to be created. The following section discusses how the system would work.

@@

Once a web application is created, there are two options to run it; one is to run it inside MTurk system using their API and the other is to develop a new web site separate from MTurk. For additional control we picked the second option and planned to separate this application from MTurk system and designed our own web site to run the application. It enables us to display the lineups to the observers with a lot of flexibilities. As an added benefit the data can be directly saved in a local database server instead of getting it from MTurk.

Figure \ref{fig:turk_work_flow} shows how the plan works. First an MTurk task is created for the workers to review and decide whether they want to do the task based on the parameters like payment amount, estimated time the task may take to finish and how hard the task seems to be etc. Workers are informed that the task has to be done outside the MTurk system from another web site. Once the workers accept the task they are redirected to our web site where multiple lineups are shown for evaluations. After the required number of evaluations has been received, a code is generated which the workers need to submit back in MTurk system to complete the task. The code is matched with the code in our database to process payment through MTurk system.

@@

If a lineup is created from the actual observed data, the web application presented in Section \ref{sec:web_application} can be used to get it evaluated. In that case we do not need to design a simulation experiment. This section will be useful when a simulation experiment is needed to examine different visual test statistics and compare the power \citep{majumder:2013}. To examine the effectiveness of certain plots in displaying the data one may want to design an experiment with lineup \citep{heike:2012}. In any cases, the two major considerations are the choice of parameters to simulate lineups and number of people needed to evaluate a lineup.

\subsection{Selecting Parameter to Simulate Lineup} \label{sec:parameter_selection} In any experimental design we need to have some control. For a simulation experiment with lineup, this refers to the parameters related to a lineup. This depends on what the purpose of the experiment is. For example, to compare the power of a parametric test one may fix the parameter specified by alternative hypothesis and generate data from that model. This data can be considered as the observed data. A lineup can be created by placing this observed data plot in a layout where rest of the data plots come from the model specified by the null model \citep{majumder:2013}.

In addition to the hypothesized parameter of interest, other parameters that may need to be fixed include sample sizes, variability in the data, error structure etc. These parameters produce the effect sizes and are responsible for any detectable signal in the lineup. Thus for a simulation study a range of effects sizes need to be considered. It is a little bit tricky to decide what exact effect sizes are appropriate. For example in \cite{majumder:2013} the parameters were chosen so that they produce a smoother power curve of a conventional test. But in general this depends on what the researchers may want to test and control in an experimental study.

@@

\subsection{Procedure to Simulate Data Plot} \label{sec:simulate_plot} In the simulation study the very first step is to generate a random sample of data from the model of interest with pre-specified parameters. We call this data set observed data. Every time we do that we get a new set of observed data which may produce an estimated parameter very different than the actual parameter specified to generate the data. But it is important to have an observed data closely representing the parameters chosen since other null plots will be compared to this data plot in a lineup. One solution to this is to take some replications of lineups for the same effect size.

While the effect of the natural variability of the observed data set can be controlled by taking the replication of few samples, it is desirable to study whether we can reduce this variability further to some extent. For this we pick the example of a simple linear regression model and examines how the data plot can be generated that best represents the specified parameter.  Consider a linear regression model 
\begin{equation} \label{eqn:turk_model} 
Y_i = \beta_0 + \beta X  + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariate $X$ can be continuous or discrete. As discussed in Section \ref{sec:parameter_selection} we specify the parameters sample size $n$=300, regression slope $\beta$=3 and error standard deviation $\sigma$=12 to simulate observed data.

@@

{\bf Quantiles of $p$-value approach:} For a simulated observed data set we can obtain the $p$-value associated with testing $H_0: \beta=0$. The distribution of $p$-value is uniform under null hypothesis but under alternative it has a right skewed distribution. Figure~\ref{fig:dist_pvalue} shows the distribution  of $p$-values for sample size $n$=300, regression slope $\beta$=3 and error standard deviation $\sigma$=12. If we generate a data set randomly there is a 21\% chance that the data will show a $p$-value of 0.25 or more even though we generated data set with non-zero $\beta=3$. We need to make sure that the simulated data does not come from this extreme end. Additionally, we want some replications so that this extreme effect can be controlled at some level. For this example we intend to pick three replications.

\begin{figure}[hbtp]
   \centering
       \scalebox{.6}{\includegraphics{dist_pvalue.pdf}}
       \caption{Distribution of $p$-values under alternative hypothesis ($H_1: \beta$=3) for sample size $n$=300 and error standard deviation $\sigma$=12.}
       \label{fig:dist_pvalue}
\end{figure}

@@

{\bf Closeness of estimated parameters:} When we simulate a data set from the Model \eqref{eqn:turk_model} and fit the model to the data set again, we do not get the parameters estimates same as what we fixed while simulation. For example, with a simulated observed data we obtained a parameter estimates of 0.5 for true $\beta=3$. We want a data set where the parameter is closed to our fixed slope $\beta=3$.  In this approach we generate 1000 data sets and pick the data set that shows most close estimates of parameters compared to true parameter values. We do this three times to obtain three replications of the data.

@@

\subsection{Sample size estimation} For any experiment it is important to determine the sample size. It is not only related to time and cost of the experiment but also associated with the validity of the results. In simulation experiment with lineup the sample size means the number of evaluations per lineup. It is different than the number of people who evaluates lineup. If multiple lineups are evaluated by a person, sample size will be larger than number of people recruited from MTurk. For example, if a lineup needs to be evaluated 20 times, we need at least 20 persons since each person does not see the same lineup more than once. This will also enable us to get 10 lineups evaluated if all of these 20 people evaluate 10 lineups each. But all the lineups don't require same number of evaluations. For easy or difficult lineups fewer evaluations are needed.

One approach to assess number of evaluations that are required for a lineup is to have a prior idea of proportion correct responses for the that lineup. For a given proportion $p$ we want to have margin of error (ME) to be at most 0.05. Thus we have $$ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$$ which gives us the estimation of minimum sample size $$n \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$

If we do not have any prior idea of proportion of correct or power, we may rely on the power of a similar conventional test if available. This can be assumed to be the power or proportion correct for that lineup. Other ways of having an estimate of sample size can be very specific to the problem of interest and should be computed as required by the specific problem.
